### 2024-11-12 submission/manuscript.Rmd
.Rmd run successfully to create .tex .pdf

---
title: "**NAME OF THIS STUDY**"
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header.tex
csl: mbio.csl
fontsize: 11pt
geometry: margin=1.0in
---

## ff
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
knitr:: opts_chunk$set("tidy" = FALSE)
knitr:: opts_chunk$set("echo" = FALSE)
knitr:: opts_chunk$set("eval" = TRUE)
knitr:: opts_chunk$set("warning" = FALSE)
knitr:: opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace{35mm}

Running title: INSERT RUNNING TITLE HERE

\vspace{35mm}

Munkh^1, Joeseph P. Schmo^2, Sally J. Rivers^1, Patrick D. Schloss^1$\dagger$^

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

1\. Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI 48109

2\. Other department contact information

\newpage
\linenumbers

## Abstract

\newpage

## Introduction

## Results and Discussion

## Conclusions

## Materials and Methods

\newpage

Insert figure legends with the first sentence in bold, for example:

**Figure 1. Number of OTUs sampled among bacterial and archaeal 16S rRNA gene sequences for different OTU definitions and level of sequencing effort.** Rarefaction curves for different OTU definitions of Bacteria (A) and Archaea (B). Rarefaction curves for the coarse environments in Table 1 for Bacteria (C) and Archaea (D).

\newpage

## References