### 2024-11-19 submission/manuscript.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  11pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Spatio-temporal distinct patterns in variations of PM\_\{10\} and PM\_\{2.5\} relative to the recent drivings of emission sources in Mongolia},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Spatio-temporal distinct patterns in variations of
\(PM_{10}\) and \(PM_{2.5}\) relative to the recent drivings of emission
sources in Mongolia}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\vspace{35mm}

Running title: INSERT RUNNING TITLE HERE

\vspace{35mm}

Munkh\^{}1, Joeseph P. Schmo\^{}2, Sally J. Rivers\^{}1, Patrick D.
Schloss\textsuperscript{1}\(\dagger\)

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

1. Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

2. Other department contact information

\newpage
\linenumbers

\subsection{Abstract}\label{abstract}

Storyline:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A new pattern is emerged
\item
  Air quality in urban sites is episodically dictated by dust events in
  spring or late autumn, yet seasonally governed by anthropogenic
  emissions in winter. {[}Air quality is governed by natural dust
  emission, and anthropogenic emissions{]}
\item
  With recent growing interest in urban life style, and combustion of
  coal/oyutolgoi for heating winter conditions results a highly increase
  in not only capital city but also towns
\item
  In a result, spring coarse dust, plus winter fine pollutants
\item
  spring coarse dust is immediately transported and deposited in the
  source area, whereas winter fine pollutants is permanently stayed in
  the source area due to stagnant atmosphere govern over entire
  country., perhaps floating in the near surface, deposits in the
  surface{]}
\item
  Alarms, the Mongolian dust in the spring, optical properties might be
  shifted; this gives \ldots{} Gobi dust and sand storms has become
  tuiren, from the shoroon shuurga. which clearly requires the
  attention.
\item
  r ratio shows \ldots{} emission source; dust might carry anthropogenic
  fine particulates as well. \newpage
\end{enumerate}

\subsection{Introduction}\label{introduction}

\subsection{Research Qs}\label{research-qs}

Therefore, we aimed to demonstrate the distinct temporal and spatial
variations of PM2.5 and PM10 across urban and rural Mongolia using
extensive data from 2008 to 2020.

On spring, the dust storm from the Gobi Desert contribute significantly
to increased aerosols in the atmosphere and ambient air pollution,
leading to sporadic peaks in PM10 concentrations reaching as high as
64-234 \(\mu g m^{-3}\) per day or exceeding 6000 \(\mu g m^{-3}\) per
hour (Jugder). concentrations of particulate matter is ephederemal, yet
vary depending on whether the pollution cause is natural or industrial,
local or transported, seasonal or non-seasonal, makes complex and
challenging. 1. Do concentrations of particulate matters differ in
between urban and rural sites, and even within Gobi sites? 2. Do
distinct temporal variations has existed among the sites? 3. Do PM2.5
particulates has contributed to the PM10 annual variations?

\begin{verbatim}
-   If yes, how much, and when and where?
-   What is the sd, mean, and median
    -   box plot
    -   violin
    -   scatter points, epidemic, sporadic
-   Daily variations to examine it related to the heating
    -   2 peaks: smaller and bigger
    -   compare the t-duration exceeds 50mug/m3/hour

4.  Does it has distinct patterns among the sites regarding to the
    drivings

-   How PMs varies with the wind speed and visibility
-   Do they differently explained with variables and changes in
    drivings (with PCA analysis)

5.  Is there any significant changes in time-series of PMs at 4
    seasons
6.  Is there any significant changes in ratio in the spring in
    respect to winter?

    The present study will contribute significantly to the understanding
        of air particulate matter patterns in Mongolia and providing
        comprehensive data insights for policymakers and public health
        sectors. Our findings is useful not only for addressing national health
        impacts but also beneficial for understanding air particulate matter
        as ambient air pollution, and tackling atmospheric aerosol effects
        in the climate system, and revealing their transboundary effects to
        the downwind regions in South-east Asia.
\end{verbatim}

\subsection{Results}\label{results}

\subsubsection{The spatio-temporal variations of the PMs at the study
sites}\label{the-spatio-temporal-variations-of-the-pms-at-the-study-sites}

To comparatively examine the spatio-difference regarding to the
emissions of PMs, we illustrated the hourly observed values of PM10 and
PM2.5 for all sites (figure\_3). On each site, PM10 is more sporadic
than PM2.5 due to dust epidemic nature. In conjunction, the mean values
averaged from hourly values are larger than its median concentrations
both on PM10 and PM2.5 for all sites. The mean variables of p-value show
that concentrations of particulate matters significantly differ with
99\% confident levels at all sites (figure\_3), except 95\% confident
level between DZ and UB on PM10 (figure\_3a). It clearly exhibits
general variations of particulate matters as expected and emphasizes
that the significant difference in the concentration values among the
sites. Besides all sites are quantitatively differ by the values of
concentrations in particulate matters, there are two main
characteristics can be seen when we compare the median deviations from
its mean values etc.

\begin{itemize}
\tightlist
\item
  such patterns are strongly manifested in ZU and SS sites.\\
\item
  the mean values those averaged from hourly concentrations of PM10 and
  PM2.5 is larger than its median are strongly manifested for all in UB
  and DZ sites.
\item
  Additionally, consider the significance of the ..
\end{itemize}

This highlight that the fine and coarse particulate matters
significantly vary among sites of urban and rural sites, and even within
Gobi (rural) sites. Moreover, \ldots. has a diversely characteristically
diversified in respect with the PM2.5 (fine) to PM10 (coarse)
particulates.,. which requires urban impacts

To reveal the natural and anthropogenic impacts on the concentrations of
particulate matters, we demonstrated annual variations of PM10 and PM2.5
for each sites (figure\_4). 3. Do PM2.5 particulates has contributed to
the PM10 annual variations? Distinct temporal variations has existed
among the sites. - PM2.5 particulates has contributed to the PM10 annual
variations in UB and in DZ in winter. - If yes, how much, and when and
where? - What is the sd, mean, and median - box plot - violin - scatter
points, epidemic, sporadic - Daily variations to examine it related to
the heating - 2 peaks: smaller and bigger - compare the t-duration
exceeds 50mug/m3/hour

{[}Therefore{]} SS and ZU sites are mainly affected by the spring dust,
followed by the autumn dust. Annual maximum in the winter for DZ and UB
are from PM2.5, which results an increase in PM10. It requires the cause
the behind such the variations. DZ site is polluted in the winter by the
heating and in the spring by the natural dust.

{[}Therefore{]} Spatio-temporally in two class; consists of 2 Gobi
sites, and 1 urban plus urbanized Gobi sites.

\subsubsection{The emission patterns of interrelations among
meteorological variables at the study
sites}\label{the-emission-patterns-of-interrelations-among-meteorological-variables-at-the-study-sites}

\begin{verbatim}
    -   PMs varies with the wind speed and visibility
    -   In general, three distinct patterns were resulted with PCA
        analysis, which is in consistent with temporal variation.
        explained with variables and changes in drivings (with PCA
        analysis)
\end{verbatim}

DZ site is polluted in the winter by the heating and in the spring by
the natural dust.

\subsubsection{The recent trends in concentrations of PMS and
fine-coarse}\label{the-recent-trends-in-concentrations-of-pms-and-fine-coarse}

\begin{verbatim}
    fractional changes at the sites
    -   There are significant changes in time-series of PMs at 4
        seasons
    -   There any significant changes in ratio in the spring in
        respect to winter in DZ.
    -   Close relationships was found between PM2.5 in winter and r
        values in the spring.
\end{verbatim}

\subsection{Conclusions}\label{conclusions}

\begin{itemize}
\tightlist
\item
  The spatio-temporal variations of the PMs at the study sites -
  Concentrations of particulate matters differ in between urban and
  rural sites, and even within Gobi sites. - Distinct temporal
  variations has existed among the sites. - PM2.5 particulates has
  contributed to the PM10 annual variations in UB and in DZ in winter. -
  If yes, how much, and when and where? - What is the sd, mean, and
  median - box plot - violin - scatter points, epidemic, sporadic -
  Daily variations to examine it related to the heating - 2 peaks:
  smaller and bigger - compare the t-duration exceeds 50mug/m3/hour

  \begin{itemize}
  \tightlist
  \item
    The emission patterns of interrelations among meteorological
    variables at the study sites

    \begin{itemize}
    \tightlist
    \item
      PMs varies with the wind speed and visibility
    \item
      In general, three distinct patterns were resulted with PCA
      analysis, which is in consistent with temporal variation.
      explained with variables and changes in drivings (with PCA
      analysis)
    \end{itemize}
  \item
    The recent trends in concentrations of PMS and fine-coarse
    fractional changes at the sites

    \begin{itemize}
    \tightlist
    \item
      There are significant changes in time-series of PMs at 4 seasons
    \item
      There any significant changes in ratio in the spring in respect to
      winter PM2.5 in DZ.
    \end{itemize}
  \item
    Close relationships was found between PM2.5 in winter and r values
    in the spring. Thus, our research results clearly proves the
    distinct variations in PMs has emerged. The dust fine-coarse
    fractions was manifested at the town center for the Gobi sites,
    which reveals the that Mongolian dust composites not only coarse
    dust, but also fine particulate matters. The particulates likely
    consisted of the black carbon, which may give a substantial effect
    on climate systems. if this trend continues on as coal consumption
    with the population growth in the future.
  \item
    CO Carbon monoxide is obtained due to incomplete combustion of
    charcoal in a closed room.
  \item
    CO2
  \end{itemize}
\end{itemize}

\newpage

\subsection{Results and Discussion}\label{results-and-discussion}

\newpage
\subsection{Urban and rural distinct Spatio-temporal diverse variations of $PM_{10}$ and $PM_{2.5}$}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_3.png}
\caption{Distinct concentrations of coarse and fine particulates among
sites}
\end{figure}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compare the concentrations of PMs at UB is the 2. Significance level
  difference 3. Conclude
\end{enumerate}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.4375in,height=\textheight,keepaspectratio]{images/figure_4.png}
\caption{Annual variations of \$PM\_\{10\}\$ and \$PM\_\{2.5\}\$}
\end{figure}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Clear annual variations at UB and DZ from pm2.5 pollutions 2. at ZU,
  and SS has a seasonally peaks episodic spring and late autumn from
  PM10
\end{enumerate}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_5.png}
\caption{Daily variations of \(PM_{10}\) and \(PM_{2.5}\) at UB and DZ
sites}
\end{figure}

\newpage
\subsection{Meteorological influence on $PM_{10}$ and $PM_{2.5}$ variations}
\label{subsec2}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_6.png}
\caption{Relationships between meteorological major factors and
variations of \(PM_{10}\) and \(PM_{2.5}\)}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7.png}
\caption{Spatio-temporal distinct feature of variations of \(PM_{10}\)
and \(PM_{2.5}\) with PCA analysis}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7b.png}
\caption{Patterns of meteorology and PMs at the 4 sites}
\end{figure}

\newpage
\subsection{Trends}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_8.png}
\caption{Interannual and seasonal trends of \(PM_{10}\) and \(PM_{2.5}\)
variations}
\end{figure}

\newpage

\subsection{Conclusions}\label{conclusions-1}

In this study, we investigated the temporal variations of PM2.5 and PM10
concentrations at the 4 sites of rural and urban those located along the
the wind corridor. Three distinct variations has been detected.

\begin{enumerate}
    \item Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
\end{enumerate}

A clear seasonal variations in the sites of UB and DZ is {[}Air quality
is governed by natural dust emission, and anthropogenic emissions{]} *
Due to rapid increase in urban, and combustion of coal/oyutolgoi for
heating winter conditions results a highly increase in not only capital
city but also towns * In a result, spring coarse dust, plus winter fine
pollutants {[}spring coarse dust is immediately transported and
deposited in the source area, whereas winter fine pollutants is
permanently stayed in the source area due to stagnant atmosphere govern
over entire country., perhaps float- ing in the near surface, deposits
in the surface{]} * Alarms, the Mongolian dust in the spring, optical
properties will be shifted; this gives \ldots{} Gobi dust and sand
storms has become tuiren, from the shoroon shuurga. which clearly
requires the attention.

Following problems

\begin{itemize}
\tightlist
\item
  On downwind regions
\item
  On national-level Demonstrating temporal and spatial variations of air
  particulate matter has become important for understanding
  characteristics of particulate matter in the climate system, providing
  valuable information for well-established air quality measures, and
  illustrating the good trace data for health studies. Because
  particulate pollutants have a great impact on human health (Dockery
  and Pope,1994; Harrison and Yin, 2000; Hong et al., 2002), high
  atmospheric concentrations of these pollutants was a major concern
  particularly in urban areas, in the last 2-3 decades. Recent studies
  highlight that even low concentrations of these pollutants can lead to
  various health issues, and may associate with morbidity and mortality
  across the life span (Zigler et al., 2017). Children exposed to high
  levels of air pollution show increased rates of asthma, decreased lung
  function growth, and increased risk of early markers of cardiovascular
  disease (Bourdrel et al., 2017; Gauderman et al., 2015; Hehua et al.,
  2017). Short-term exposure with high level of PM10 resulted the
  chronic cardiovascular disease in Mongolia (Enkhjargal 2020). In
  addition to these health issues, (prenatal) neurodevelopmental impacts
  such as effects on intelligence, attention, autism, and mood, while
  aging populations experience accelerated cognitive decline when
  exposed to high levels of pollution is detected (Power et al., 2016).
  Long-term exposure to low levels of particulate matter, such as
  concentrations as low as 10 \(\mu g m^{-3}\) (equilibrium to WHO Air
  Quality Guidelines), has been linked to increased lung cancer in the
  EU (Hvidtfeldt et al. 2021), with similar evidences reported in Canada
  (Bai et al., 2019), and significantly higher rates captured in China
  with concentrations up to 30 \(\mu g m^{-3}\). Apparently, pollutants
  of particulate matters has effects to various health issues with the
  different thresholds and exposure durations. However, more in-depth
  and diversified research on air pollution and its health effects is
  essential, with the detailed information is necessary (Tan et al

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi})}
  \setcounter{enumi}{2020}
  \tightlist
  \item
    to have accuracy of assessing exposure to air pollution during
    developmentally relevant time periods, such as trimesters or months
    (Becerra et al., 2013; Gong et al., 2014; Kalkbrenner et al., 2014)
    or weeks (Chiu et al., 2016). Many research findings/Numerous
    research findings have advanced the field, and air quality indices
    is widely used for providing guidance, and public perception of air
    quality has been improved (Mirabelli et al., 2020).
  \end{enumerate}
\end{itemize}

\newpage

\subsection{Materials and Methods}\label{materials-and-methods}

\subsubsection{A description of study
sites}\label{a-description-of-study-sites}

According to the spatial magnitude of wind stress in Mongolia (Figure
1), the largest magnitude of wind speed is on the Gobi sites,
particularly those located in the southeast edge of the country.

\begin{itemize}
\tightlist
\item
  The impact of high winds on plant diversity varies across
  environmental gradients of precipitation and soil fertility (Milchunas
  et al., 1988).
\item
  In the desert steppe zone, species richness was lower in the drier
  years but did not vary with grazing pressure.
\item
  In the steppe zone, species richness varied significantly with grazing
  pressure but did not vary between years. Species richness is not
  impacted by grazing gradient in desert steppe, but it is in the steppe
  (Cheng et al., 2011).
\end{itemize}

In the last 2 decades, due to poverty and natural disasters there is
population immigration has taken place from the rural to urban,
especially to capital city of Mongolia. Due to tiny infrastructure to
provide the mega city with the dense population, it introduces the urban
pollution. Therefore, Ulaanbaatar air particulate matter mainly reflects
the coal burning, and partly, natural dust.

Consequently, the atmospheric environment and climate for Mongolian Gobi
has been impacted the most by frequent dust and and sand storm in the
spring.

Our study was carried out in Dalanzadgad (town center) (Tbl. 1; 43.57°N,
104.42°E), Sainshand (Tbl. 1; 44.87°N, 110.12°E) and Zamyn-Uud (Tbl. 1;
43.72°N, 111.90°E) in the Gobi Desert, and at Ulaanbaatar (Tbl.??.??°N,
104.42°E) (city center) located in the temperate Mongolian steppe of
Mongolia (Figure 2). Nomads and settlements of this sum have raised a
large number of livestock, and they rank at number 30 out of 329 sums
for the largest number of livestock raised per sum (Saizen et al.,
2010). In the last decade, the number of dust events associated with
wind erodibility increased by 30 \% in Bayan-Önjüül (Kurosaki et al.,
2011). This is an area where dust emissions activity has been monitored
on a long-term basis (Shinoda et al., 2010a) at a dust observation site
(DOS) adjacent to the study site (Fig. 1a). According to long-term
meteorological observations made at the monitoring station of the
Institute of Meteorology and Hydrology of Mongolia located near the
site, the prevailing wind direction is northwest. Mean annual
precipitation is 163 mm, and mean temperature is 0.1◦C for the period
1995 to 2005 (Shinoda et al., 2010b). Soil texture is dominated by sand
(98.1 \%, with only 1.3 \% clay and 0.6 \% silt; Table 1; Shinoda et
al., 2010a). Insert figure legends with the first sentence in bold, for
example:

\newpage

\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_1.png}{]}

\begin{figure}
\centering
\includegraphics[width=4.16667in,height=\textheight,keepaspectratio]{images/table_1.png}
\caption{\textbf{Table 1}. A description of datasets obtained at the
sites}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/scheme_1.png}
\caption{Scheme 1. Data handling procedure}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2b.png}
\caption{Figure 2. Data gap filling}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2c.png}
\caption{Figure 2b. Data gap filling}
\end{figure}

\newpage

\subsection{References}\label{references}

\end{document}