### 2024-11-22 submission/Outline_20241121.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  11pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Spatio-temporal distinct patterns in variations of PM\_\{10\} and PM\_\{2.5\} relative to the recent drivings of emission sources in Mongolia},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Spatio-temporal distinct patterns in variations of
\(PM_{10}\) and \(PM_{2.5}\) relative to the recent drivings of emission
sources in Mongolia}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\vspace{35mm}

Running title: INSERT RUNNING TITLE HERE

\vspace{35mm}

Munkhtsetseg, Shimizu. A, Matsuki. A, Batdorj. D, Matsui. H

\vspace{40mm}

To whom correspondence should be addressed:

\newpage
\linenumbers

\subsection{Abstract (150 words)}\label{abstract-150-words}

Storyline:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A new pattern is emerged
\item
  Air quality in urban sites is episodically dictated by dust events in
  spring or late autumn, yet seasonally governed by anthropogenic
  emissions in winter. {[}Air quality is governed by natural dust
  emission, and anthropogenic emissions{]}
\item
  With recent growing interest in urban life style, and combustion of
  coal/oyutolgoi for heating winter conditions results a highly increase
  in not only capital city but also towns
\item
  In a result, spring coarse dust, plus winter fine pollutants
\item
  spring coarse dust is immediately transported and deposited in the
  source area, whereas winter fine pollutants is permanently stayed in
  the source area due to stagnant atmosphere govern over entire
  country., perhaps floating in the near surface, deposits in the
  surface{]}
\item
  Alarms, the Mongolian dust in the spring, optical properties might be
  shifted; this gives \ldots{} Gobi dust and sand storms has become
  tuiren, from the shoroon shuurga. which clearly requires the
  attention.
\item
  r ratio shows \ldots{} emission source; dust might carry anthropogenic
  fine particulates as well.
\end{enumerate}

\newpage

\subsection{Introduction}\label{introduction}

\begin{itemize}
\item
  Advanced the knowledge of global dust, has reached to recognize the
  sources,.
\item
  Classification dust brown color, seasonal characteristics, with coarse
  fractions.
\item
  This knowledge further efficient to climate system when elaborating
  dust-aerosol effects.
\item
  But, a large uncertainties in the global dust model has existed so for
  climate models which clearly limits our understanding the climate
  system and shape the facing global issues of global warming.
\item
  This is mainly caused by the lack of parameterization and recognition
  of iterative changes controlled by the natural forces and
  anthropogenic drivings.
\item
  Mongolian dust brown color, seasonal characteristics, with coarse
  fractions.
\item
  Mongolian dust has an attention of the its mass fraction in global
  dust, yet unlikely elaborated in the climate models due to its
  majority of coarse fraction for its a small contribution to the
  climate system through its radiative feedback.
\item
  But, such recognized characterization might get no longer valid due to
  recent change in the driving of the emissions of air particulate
  matters. A large high concentrations of PM2.5 in the capital city of
  Mongolia has been observed as a result of the heavy consumptions of
  coal as a winter heating has rapidly spread as a mining industry taken
  off since 2000. Winter weather stagnant conditions governed by the
  Siberian magnifies the concentrations of the particulate matter
  emissions by trapping the polluted air below the boundary layer, so
  that results in a very large high concentrations of PM2.5, locally.
  Even recognized as one of the highly polluted capital cities in the
  world.
\item
  Therefore, It is important to examine the emerging changes and
  shifting patterns of air particulate matters in Mongolia. More
  importantly, it is essential to reveal the significant changes in the
  the altered fraction particularly, in the dust seasons considering its
  high potential of intriuging in the free atmosphere to transported in
  the long-distance, so carrying capacity of the role to shift the
  global climate system, and its side impacts on downwind regions.
\item
  Study goal - We hypothesize \ldots{} - Our study will benefit not only
  to the global dust research but also climate, and further to the
  country itself for urban planning, and coal combustion.
\end{itemize}

\subsubsection{Research Qs}\label{research-qs}

Therefore, we aimed to demonstrate the distinct temporal and spatial
variations of PM2.5 and PM10 across urban and rural Mongolia using
extensive data from 2008 to 2020.

On spring, the dust storm from the Gobi Desert contribute significantly
to increased aerosols in the atmosphere and ambient air pollution,
leading to sporadic peaks in PM10 concentrations reaching as high as
64-234 \(\mu g m^{-3}\) per day or exceeding 6000 \(\mu g m^{-3}\) per
hour (Jugder). concentrations of particulate matter is ephederemal, yet
vary depending on whether the pollution cause is natural or industrial,
local or transported, seasonal or non-seasonal, makes complex and
challenging. 1. Do concentrations of particulate matters differ in
between urban and rural sites, and even within Gobi sites? 2. Do
distinct temporal variations has existed among the sites? 3. Do PM2.5
particulates has contributed to the PM10 annual variations?

\begin{itemize}
\tightlist
\item
  If yes, how much, and when and where?
\item
  What is the sd, mean, and median

  \begin{itemize}
  \tightlist
  \item
    box plot
  \item
    violin
  \item
    scatter points, epidemic, sporadic
  \end{itemize}
\item
  Daily variations to examine it related to the heating

  \begin{itemize}
  \tightlist
  \item
    2 peaks: smaller and bigger
  \item
    compare the t-duration exceeds 50mug/m3/hour
  \end{itemize}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Does it has distinct patterns among the sites regarding to the
  drivings
\end{enumerate}

\begin{itemize}
\tightlist
\item
  How PMs varies with the wind speed and visibility
\item
  Do they differently explained with variables and changes in drivings
  (with PCA analysis)
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Is there any significant changes in time-series of PMs at 4 seasons
\item
  Is there any significant changes in ratio in the spring in respect to
  winter?
\end{enumerate}

The present study will contribute significantly to the understanding of
air particulate matter patterns in Mongolia and providing comprehensive
data insights for policymakers and public health sectors. Our findings
is useful not only for addressing national health impacts but also
beneficial for understanding air particulate matter as ambient air
pollution, and tackling atmospheric aerosol effects in the climate
system, and revealing their transboundary effects to the downwind
regions in South-east Asia.

\subsection{Results}\label{results}

\subsubsection{The spatio-temporal variations of the PMs at the study
sites}\label{the-spatio-temporal-variations-of-the-pms-at-the-study-sites}

To evaluate the spatial variations in particulate matter (PM)
concentrations, we displayed hourly observed values of PM10 and PM2.5
for all study sites (figure\_3). The mean p-values indicate that PM
concentrations differ significantly at a 99\% confidence level across
all sites (figure\_3), with the exception of a 95\% confidence level
between DZ and UB for PM10 (figure\_3a), highlighting substantial
concentration disparities among sites. While quantitative differences in
PM concentration values exist across all sites, two key patterns emerge
when examining median deviations from mean values and irregular
observation fluctuations. For instance, PM10 demonstrates more erratic
behavior than PM2.5 at each location, particularly evident at ZU and SS
sites. Furthermore, the mean values calculated from hourly measurements
surpass the median concentrations for both PM10 and PM2.5 across all
sites, with notable prominence at UB and DZ locations. Consequently,
significant spatial differences in PM concentrations exist among all
sites, regardless of urban or rural classification. However, the sites
can be categorized into two groups based on their characteristics: UB
(urban) and DZ (rural town, Gobi); and SS (rural, Gobi) and ZU (rural,
Gobi). These findings for DZ appear to support our hypothesis of
emerging new emission patterns related to increased coal consumption
during winter months.

--\textgreater{} 3.1.2

To investigate whether emerging PM patterns are associated with
household winter heating activities, we demonstrated annual variations
in PM10 and PM2.5 concentrations at the sites. Significant annual
variations in PM10 and PM2.5 levels were observed at UB and DZ sites,
with maximum concentrations exceeding 100 \(\mu g/m^3\) during colder
months (January, November, December) and lower levels consistently below
50 \(\mu g/m^3\) during warmer months (May-September). These annual
maximums coincided with the diurnal variations in PM10 and PM2.5
concentrations at sites DZ and UB, where PM concentrations reached their
highest values during nighttime and early morning hours (approximately 8
PM to 4 AM UTC), with median values surpassing 50 \(\mu g/m^3\).
Conversely, both pollutants exhibited reduced concentrations during
daytime hours (8AM to 4PM UTC), likely due to increased atmospheric
dispersion. UB site exhibited similar daily fluctuations with extended
periods (approximately 8 PM to 5 AM UTC) of elevated concentrations.
Additionally, winter PM10 concentrations at both sites were primarily
composed of PM2.5 (figure\_5; mean values for PM10 with the color bar).
The increase in PM10 and PM2.5 aligns with the heating active hours,
suggesting that household coal consumption contributes to elevated PM
levels at both DZ and UB sites. In contrast, ZU and SS sites displayed
significantly lower annual PM10 and PM2.5 levels, with sudden frequent
spikes in spring followed by occasional instances in autumn. These
annual variations were more pronounced in PM10 compared to PM2.5,
highlighting the impact of Gobi dust and sand storms. Similar
occurrences were also noted at the DZ site, suggesting its exposure to
both winter heating emissions and natural spring dust, reflecting its
Gobi-region characteristics. The annual variability in PMs with higher
concentrations during nighttime and colder months, indicating the
influence of localized emission sources and reduced boundary layer
mixing at UB and DZ sites. Additionally, the upward extended ranges
without the bottom bottle of the violin plot, demonstrating greater
variability during colder months. It implies instability in
concentrations potentially reaching high levels above 400 \(\mu g/m^3\)
when Arctic oscillation/Siberian high intensifies with heating, and
dropping below 50 \(\mu g/m^3\) when it weakens or heating demand
decreases. These findings confirm the (DZ) emerging PM pattern is caused
by household activities and influenced by meteorological conditions.
Overall, meteorological factors appear to play a crucial role in
governing PM levels.

--\textgreater{} 3.2.1

\subsubsection{The emission patterns of interrelations among
meteorological variables at the study
sites}\label{the-emission-patterns-of-interrelations-among-meteorological-variables-at-the-study-sites}

\begin{itemize}
\tightlist
\item
  PMs varies with the wind speed and visibility
\item
  In general, three distinct patterns were resulted with PCA analysis,
  which is in consistent with temporal variation. explained with
  variables and changes in drivings (with PCA analysis)
\end{itemize}

{[}AND{]} To distinguish the emission driving variables, we demonstrated
interrelations between(WS) and visibility (VIS) among particulate
matters of PM10 and PM2.5 for study sites (figure\_4). {[}AND{]} Higher
PM concentrations tend to correspond to lower wind speeds (smaller
circles) and reduced visibility (darker blue points) in UB. {[}AND{]}
All Gobi sites show the largest variability in PM concentrations, with
extreme outliers correlated with the WS primarily in the springs.

{[}AND{]} In comparison between PM10 increase with the PM2.5, as
concentrations increase in PM10 concentration the ratio 1:1 is
inscreased approximately 1.3 times for UB, and 2 times for all three
Gobi sites. This increase in Gobi sites were further increased upto 5
times in ZU, and XXX times in SS sites showing the insufficient PM2.5
particulates in the those sites.

The Principal Component Analysis (PCA) biplot (Figure 7a) highlights the
relationships among key variables, including PM10, PM2.5, wind speed
(WS), wind direction (WD), visibility, relative humidity (r), and
optical properties of aerosols (OPC). The first two principal components
(Dim 1 and Dim 2) explain 66.52\% of the total variance, with Dim 1
(44.36\%) strongly associated with PM10 and PM2.5, indicating that these
pollutants are major contributors to air quality variations. Visibility
is inversely correlated with PM10 and PM2.5, suggesting that reduced
visibility aligns with higher pollution levels. Wind speed and wind
direction are positively aligned with Dim 2 (22.16\%), reflecting their
influence on pollutant dispersion.

The clustering of monitoring sites based on PM concentrations and
geographic characteristics (Figure 7b) reveals distinct patterns. UB
(urban background) is characterized by high PM concentrations (positive
Dim 1 scores), while SS (suburban site) demonstrates low PM levels,
clustering tightly along the negative Dim 1 axis. DZ (industrial site)
exhibits substantial variability, with clusters extending into higher
Dim 1 and Dim 2 ranges, reflecting the complex interplay of emission
sources and meteorological factors. ZU (rural site) overlaps with SS but
shows greater spread, indicating moderate pollution levels influenced by
seasonal and localized factors. These results underscore the significant
spatial and meteorological influences on air quality across the studied
locations.

{[}BUT{]}

{[}THEREFORE{]} 3 clusters, supply is different?

\subsubsection{The recent trends in concentrations of PMS and
fine-coarse fractional changes at the
sites}\label{the-recent-trends-in-concentrations-of-pms-and-fine-coarse-fractional-changes-at-the-sites}

Figure 8 illustrates the trend analysis of PM10 and PM2.5 concentrations
across four monitoring sites (UB, DZ, ZU, and SS) from 2009 to 2020. The
time series demonstrates seasonal patterns and trends, with notable
differences in particulate matter concentrations between the four
seasons: Q1 (winter), Q2 (spring), Q3 (summer), and Q4 (autumn).

The p-values displayed on the figure indicate the statistical
significance of trend changes in PM10 and PM2.5 levels for each season.
At UB, significant negative trends are observed for Q1 and Q2 (p
\textless{} 0.001). At DZ, a significant positive trend is observed for
Q1 (R = 0.17, p \textless{} 0.001), indicating increasing particulate
matter levels during this season.17, p \textless{} 0.001), while Q2
exhibits no significant trend (R = 0, p = 0.915)., highlighting a
decreasing pattern in particulate concentrations during these periods
over time. In contrast, Q3 and Q4 show weaker or nonsignificant trends
(e.g., R = -0.03, p \textgreater{} 0.2), suggesting minimal changes in
PM levels during the warmer months.

The industrial site (DZ) exhibits higher variability in PM levels, with
periodic spikes indicative of localized emissions and industrial
activities. Urban background (UB) consistently records elevated PM
levels during Q1, driven by heating and combustion-related sources.
Rural (ZU) and suburban (SS) sites generally display lower particulate
concentrations. However, at ZU, a significant negative trend is observed
in both PM10 and PM2.5 for Q1, suggesting a consistent reduction in
particulate matter during winter months (may reduced transboundary
traffic). At SS, a negative trend is observed in PM2.5 for Q2,
reflecting a seasonal decline potentially linked to reduced wind speed
during spring.

The significant decreasing trends in PM concentrations in UB during
colder months likely indicate improvements in emission control measures
and a transition to more efficient heating practices. In contrast, a
significant increasing trend is observed at DZ, particularly during Q1,
suggesting rising emissions potentially associated with urbanized
household practices and other local activities in Mongolia's regional
towns.

\begin{itemize}
\tightlist
\item
  There are significant changes in time-series of PMs at 4 seasons
\item
  There any significant changes in ratio in the spring in respect to
  winter in DZ.
\item
  Close relationships was found between PM2.5 in winter and r values in
  the spring.
\end{itemize}

\subsection{Conclusions}\label{conclusions}

In this study, we investigated the temporal variations of PM2.5 and PM10
concentrations at the 4 sites of rural and urban those located along the
the wind corridor. Three distinct variations has been detected.

Air quality in urban sites is episodically dictated by dust events in
spring or late autumn, yet seasonally governed by anthropogenic
emissions in winter. Air quality in rural sites of SS and ZU is
episodically dictated by dust events in spring or late autumn. Air
quality in rural sites of SS and ZU is episodically dictated by dust
events in spring or late autumn.

A clear seasonal variations in the sites of UB and DZ is {[}Air quality
is governed by natural dust emission, and anthropogenic emissions{]}

\begin{itemize}
\tightlist
\item
  Due to rapid increase in urban, and combustion of coal/oyutolgoi for
  heating winter conditions results a highly increase in not only
  capital city but also towns
\item
  In a result, spring coarse dust, plus winter fine pollutants {[}spring
  coarse dust is immediately transported and deposited in the source
  area, whereas winter fine pollutants is permanently stayed in the
  source area due to stagnant atmosphere govern over entire country.,
  perhaps float- ing in the near surface, deposits in the surface{]}
\item
  Alarms, the Mongolian dust in the spring, optical properties will be
  shifted; this gives \ldots{} Gobi dust and sand storms has become
  tuiren, from the shoroon shuurga. which clearly requires the
  attention.
\end{itemize}

Following problems

\begin{itemize}
\tightlist
\item
  On downwind regions
\item
  On national-level Demonstrating temporal and spatial variations of
\end{itemize}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_3.png}
\caption{Distinct concentrations of coarse and fine particulates among
sites}
\end{figure}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compare the concentrations of PMs at UB is the 2. Significance level
  difference 3. Conclude
\end{enumerate}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.4375in,height=\textheight,keepaspectratio]{images/figure_4.png}
\caption{Annual variations of \$PM\_\{10\}\$ and \$PM\_\{2.5\}\$}
\end{figure}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Clear annual variations at UB and DZ from pm2.5 pollutions 2. at ZU,
  and SS has a seasonally peaks episodic spring and late autumn from
  PM10
\end{enumerate}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_5.png}
\caption{Daily variations of \(PM_{10}\) and \(PM_{2.5}\) at UB and DZ
sites}
\end{figure}

\newpage
\subsection{Meteorological influence on $PM_{10}$ and $PM_{2.5}$ variations}
\label{subsec2}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_6.png}
\caption{Relationships between meteorological major factors and
variations of \(PM_{10}\) and \(PM_{2.5}\)}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7.png}
\caption{Spatio-temporal distinct feature of variations of \(PM_{10}\)
and \(PM_{2.5}\) with PCA analysis}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7b.png}
\caption{Patterns of meteorology and PMs at the 4 sites}
\end{figure}

\newpage
\subsection{Trends}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_8.png}
\caption{Interannual and seasonal trends of \(PM_{10}\) and \(PM_{2.5}\)
variations}
\end{figure}

\newpage

\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_1.png}{]}

\begin{figure}
\centering
\includegraphics[width=4.16667in,height=\textheight,keepaspectratio]{images/table_1.png}
\caption{\textbf{Table 1}. A description of datasets obtained at the
sites}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/scheme_1.png}
\caption{Scheme 1. Data handling procedure}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2b.png}
\caption{Figure 2. Data gap filling}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2c.png}
\caption{Figure 2b. Data gap filling}
\end{figure}

(1)

\newpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-Munkh2017}
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Munkhtsetseg E}, \textbf{Shinoda M},
\textbf{Ishizuka M}, \textbf{Mikami M}, \textbf{Kimura R},
\textbf{Nikolich G}. 2017. Anthropogenic dust emissions due to livestock
trampling in a mongolian temperate grassland. Atmospheric Chemistry and
Physics \textbf{17}:11389--11401.
doi:\href{https://doi.org/10.5194/acp-17-11389-2017}{10.5194/acp-17-11389-2017}.}

\end{CSLReferences}

\end{document}